Clustering feature points

### 2013-05-29 text/clustering.tex
Added the MM algorithm to the clustering.tex

To get around this limitation several methods have been proposed in the 
literature. One of the most fruitful approaches have been to look at the 
geometric configuration of the feature points and match them with 
respect to assumptions concerning the variance between images. For 
example an angular and distance constraint between matches as considered 
by \cite{kim2008efficient} performs well in situations where no camera 
rotation occurs between the two images but will likely fail in other 
cases. Various scenarios have been proposed to improve on this simple 
assumption such as epipolar constraints (\cite{torr2000mlesac}, 
\cite{chum2005matching}) and pairwise constraints 
(\cite{choi2009robust}, \cite{leordenanu2005spectral}). The Epipolar 
constraint carries the assumption that the two images matched are 
related by an affine transformation. That is, there is no relative 
movement of objects in between images and either the viewpoint is fixed 
or the image resides entirely on a plane. In practice this assumption 
largely holds true when all objects we are interested in matching are 
roughly the same distance from the camera and when we don't expect to 
match objects that aren't consistently positioned across images.  
Pairwise constraints provide a more robust approach to this problem by 
looking at a set of proposed correspondences and defining a pairwise 
error between any two matches usually based on the assumption that two 
neighboring correspondences will usually have similar angles and 
distances. We can then convert the problem to an optimization problem 
and return a set of correspondences that minimizes this error such as 
proposed in \cite{choi2009robust} and \cite{leordeanu2005spectral}.  
This approach provides more robustness because assumptions that might 
not be true globally often still holds on a local level.

@@

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing the 
every feature point of one image with all feature points of the other 
and finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. By ranking the 
scores by their uniqueness and picking the $n$ best we get a set of 
correspondences that are distinctly matched across the two images. In 
the book Multiple View Geometry \cite[p.  114]{multipleView}, the gold 
standard algorithm is introduced which ranks correspondences by 
similarity and further filters them by making sure that the map of 
matches is symmetric, i.e. that for every feature point $f_i$ matching 
$f_j$ we also have that $f_j$ matches $f_i$.

@@

Based on this idea the two proposed methods find reliable matches as 
follows:
\begin{itemize}
\item{\emph{Mirror Match}: Match features using the gold standard 
algorithm\cite[p. 114]{multipleView} ranked by similarity and thresheld 
by uniqueness\cite{lowe2004sift}. However instead of matching features 
from one image with features in another, we match every feature with all 
other features of the two images combined. Only matches from one image 
to the other are returned.}
\item{\emph{Mirror Match with Clustering}: Take the combined set of 
feature points from both images and cluster these points according to 
their descriptors. Given a resulting partition of points, no matches are 
returned if it contains only feature points from one image. If the 
partition contains points from both images, \emph{Mirror Match} is used 
to find the best matches within the partition.
\end{itemize}

@@

Often images will contain repetitive patterns that are difficult to 
match\footnote{In fact this particular problem has been given attention 
before by for example \cite{fan2011towards}} because the feature points 
covering these patterns will look very similar. If we threshold by 
uniqueness like in Mirror Match this means that these points will often 
be discarded even if they might have good correspondences. To solve this 
problem the enhanced Mirror Match with Clustering makes it possible to 
look at groups of similar feature points one at a time and within each 
group finding the best matches. This is done by taking the set of 
feature points from both images and clustering. The resulting partitions 
will each contain feature points that are similar to each other. In some 
cases the clustering will end up grouping feature points of only one 
image together which can then easily be discarded in the matching 
process. In other cases partitions will contain feature points from both 
images that can then be matched using Mirror Match but with lower 
thresholds.

\subsection{Graph Clustering}

>>>>

To get around this limitation several methods have been proposed in the 
literature. One of the most fruitful approaches have been to look at the 
geometric configuration of the feature points and match them with 
respect to assumptions concerning the variance between images. For 
example an angular and distance constraint between matches as considered 
by \cite{kim2008efficient} performs well in situations where no camera 
rotation occurs between the two images but will likely fail in other 
cases. Various scenarios have been proposed to improve on this simple 
assumption such as epipolar constraints (\cite{torr2000mlesac}, 
\cite{chum2005matching}) and pairwise constraints 
(\cite{choi2009robust}, \cite{leordeanu2005spectral}). The Epipolar 
constraint carries the assumption that the two images matched are 
related by an affine transformation. That is, there is no relative 
movement of objects in between images and either the viewpoint is fixed 
or the image resides entirely on a plane. In practice this assumption 
largely holds true when all objects we are interested in matching are 
roughly the same distance from the camera and when we don't expect to 
match objects that aren't consistently positioned across images.  
Pairwise constraints provide a more robust approach to this problem by 
looking at a set of proposed correspondences and defining a pairwise 
error between any two matches usually based on the assumption that two 
neighboring correspondences will usually have similar angles and 
distances. We can then convert the problem to an optimization problem 
and return a set of correspondences that minimizes this error such as 
proposed in \cite{choi2009robust} and \cite{leordeanu2005spectral}.  
This approach provides more robustness because assumptions that might 
not be true globally often still holds on a local level.

@@

There are a relatively small number of algorithms proposed to solve the 
matching problem without involving geometric constraints. Traditionally 
the feature points of two images have been matched by comparing the 
every feature point of one image with all feature points of the other 
and finding the best matches based on the similarity of the descriptors.  
With the introduction of the SIFT features \cite{lowe2004sift}, Lowe 
proposes an alternative measure where the uniqueness of a given match is 
found by looking at the two nearest neighbors of each feature point and 
calculating a match score by the ratio of similarities. By ranking the 
scores by their uniqueness and picking the $n$ best we get a set of 
correspondences that are distinctly matched across the two images. In 
the book Multiple View Geometry \cite[p.  114]{multipleView}, the gold 
standard algorithm is introduced which ranks correspondences by 
similarity and further filters them by discarding matches that aren't 
symmetric, i.e. that for every feature point $f_i$ matching $f_j$ we 
also have that $f_j$ matches $f_i$.

@@

Based on this idea the two proposed methods find reliable matches as 
follows:
\begin{itemize}
\item{\emph{Mirror Match (MM)}: Match features using the gold standard 
algorithm\cite[p. 114]{multipleView} ranked by similarity and thresheld 
by uniqueness\cite{lowe2004sift}. However instead of matching features 
from one image with features in another, we match every feature with all 
other features of the two images combined. Only matches from one image 
to the other are returned.}
\item{\emph{Mirror Match with Clustering (MMC)}: Take the combined set 
of feature points from both images and cluster these points according to 
their descriptors. Given a resulting partition of points, no matches are 
returned if it contains only feature points from one image. If the 
partition contains points from both images, \emph{Mirror Match} is used 
to find the best matches within the partition.}
\end{itemize}

@@

Often images will contain repetitive patterns that are difficult to 
match\footnote{In fact this particular problem has been given attention 
before by for example \cite{fan2011towards}} because the feature points 
covering these patterns will look very similar. If we threshold by 
uniqueness like in Mirror Match this means that these points will often 
be discarded even if they might have good correspondences. To solve this 
problem the enhanced Mirror Match with Clustering makes it possible to 
look at groups of similar feature points one at a time and within each 
group finding the best matches. This is done by taking the set of 
feature points from both images and clustering. The resulting partitions 
will each contain feature points that are similar to each other. In some 
cases the clustering will end up grouping feature points of only one 
image together which can then easily be discarded in the matching 
process. In other cases partitions will contain feature points from both 
images that can then be matched using \emph{MM} but with lower 
thresholds.

The report is organized as follows. In section \ref{algorithms} I will 
describe the two algorithms. Section \ref{experiment} will cover the 
experimental set up and introduce the image set that is used in testing.  
In section \ref{results} I will go through the results and conclude.

\section{Algorithms}
\label{algorithms}

\begin{algorithm}
\caption{Mirror Match Algorithm}
\label{alg-mm}
\begin{algorithmic}
\Require $images$ : set of images, $t$ : float $\in \mathbb{R}$
\State $matches_{init}\gets \varnothing$
\State $matches_{final}\gets \varnothing$
\State $features\gets \varnothing$
\ForAll{$I_i \in images$} \Comment Gather features of all images
	\State $features\gets features \cup getFeatures(I_i)$
\EndFor
\ForAll{$f_i \in features$} \Comment Get best match for all features
	\State $f_m,f_n \gets getTwoNearestNeighbors(features \ {f_i})$
	\State $ratio \gets getSimilarity(f_i, f_n) / getSimilarity(f_i, 
f_m)$
	\If{$ratio < t$}
		\State $matches_{init}\gets \left(f_i, f_m\right)$
	\EndIf
\EndFor
\ForAll{$\left(f_i, f_j \right) \in matches_{init}$} \Comment Filter 
matches
\If{$\left(f_j, f_i \right) \in matches_{init} \wedge getImg(f_i) \neq 
getImg(f_j)$}
		\State $matches_{final} \gets (f_i, f_j)$
		\State $matches_{init} \gets matches_{init} \ (f_j, f_i)$
	\EndIf
\EndFor \\
\Return $matches_{final}$
\end{algorithmic}
\end{algorithm}

In this report I propose two algorithms for better matching feature 
points without any geometric assumptions about the images matched. The 
two algorithms are complementary in the sense that the first, Mirror 
Match (\emph{MM}) is less complex but performs slightly worse than 
Mirror Match with Clustering (\emph{MMC}). In cases where high

\section{Experiments and Image Set}
\label{experiment}

\section{Results}
\label{results}
\subsection{Graph Clustering}
